{\it Maxwell's Silver Hammer:}\\ On pulsed microwave acoustic resonant viral inactivation

### 2021-03-01 documents/brillouin_pulse_math.tex
Tersenating again

The extremely high losses from such thin coaxial cables should be considered.

\pagebreak
\section{Dispersion Timothy and the Brillouin Precursors}

Uzunoglu \cite{Theoretical2020} mention a handy way to circumvent this lossy-flesh problem: the Brillouin and Sommerfeld precursors\footnotemark, and pulse or wavepacket dispersion in tissues in general.\footnotemark

\footnotetext{
	Sincere thanks go to the legendary Kurt Oughstun for taking the time to reply to our emails.
}

@@

The original text of these papers are in German; translations can be found in Léon Brillouin's 1960 
book\cite{Wave1960} on the topic. \footnotemark

@@

\begin{itemize}
	\item The highly dispersive nature of tissue, which causes a peculiar distortion of pulses with broadband spectral content,
	\item The variation in loss (the imaginary part of the complex permittivity and the penetration depth) versus frequency. \wikinote{add image from that mathematica code}
\end{itemize}

A third aspect leads to the formation of the prototypical Brillouin precursor specifically: any sharp 
beginning or end of a tone, or other sharply defined structure in a waveform, includes harmonics beyond the tone's carrier wave up to the rise time \footnotemark. \wikinote{add image}

@@

Precursors have been used in practice, by Ong \cite{Detection2003} using simple square waves to detect respiration\footnotemark, Ossberger \cite{Noninvasive2004} for the same purpose; and various techniques for breast cancer detection.

Also another really good paper that I can't find.

\footnotetext{Well, Ong don't really observe the precursor itself, it seems.}

By putting more power in higher frequencies,

While it is clearly possible to produce this abnormal penetration (any sharp edge sufficiently distant from  suffices to produce a dispersive pulse), it was not obvious to us that a waveform 
could be constructed that would be an effective driver for the normal modes in question.

First, Oughston ; if the pulses are too close together, they interfere as they propagate and distort the precursor thus formed.

As an example, despite the waveform from nonlinear transmission lines producing strong wideband components, in cursory simulations they did not appear to satisfy the requirements for Brillouin pulses\wikinote{dubious-discuss}\wikinote{citation-needed}. Our working hypothesis is that this is because of destructive interference from the falling edge that immediately follows; but this is mere speculation.\wikinote{dubious-discuss}

Second, while the pulse amplitude decays slowly, the total energy contained in such a precursor decays very quickly. The biological ramfications of discussed in detail by Adair in \cite{Biophysics2000}.

temporal Soliton-producing nonlinear transmission lines, and NLTL oscillators constructed therefrom, particularly well suited

\footnotetext{it is interesting to note that a similar - if not equivalent - effect leads to the Heisenberg uncertainty principle. The more localized a }

At a depth of 8 cm in Cole-Cole muscle, an 8 GHz resonance, attack pattern alpha appears to achieve 4 picometers of oscillation amplitude, rather than 0.01 picometers with $\sin(t\ \omega_{resonance})$. The sharp edges are at approx. 250 GHz.

At 9 GHz, assuming that a torso is about 10 penetration depths in radius and the Brillouin train 
drives the virus with an equally, implementing this would decrease the field at the skin required 
for complete eradication of the virus from the body to a perfectly practicable 1 kV/m, rather than 
an air-ionizing 6.6 MV/m ($\frac{1}{\sqrt{10}}=0.316$ versus $e^{-10}=4.5e^{-5}$).

@@

Note: this isn't truly transient - the fourier integral is still periodic, wrapping around every x cycles. It's just more transient than the steady-state.

attack pattern alpha

Transient solution

@@

\begin{autem}
Reflection from tissue interfaces?

@@

Progress was hindered by the ~6 conventions for Fourier normalization, the sign or handedness of the imaginary number. What self-respecting field has completely antipodal definitions of the same concept.

oh, on the same topic, can we finally settle the i / j convention. Also, everyone that doesn't explicitly note

@@

The most promising analytical technique that we explored was a slight modification of a formulation for optimal-energy pulses, largely supplied by Prof. Costas Sarris in a yet-unpublished personal communication, based on Zhu, Hum, Costas formulation in \cite{Microwave2012b}, previously discussed by Pozar \cite{Waveform2003}.

@@

Here, rather than taking the derivative with respect to a small change in a variable, we take the derivative with respect to a small change in the function itself - a very useful technique (that the author was not aware of) which imposes little to no limitation on the form of the solution. For some intuition, the definition of the first variation and the Gateaux derivative may be of some use.

A common setup for such a problem is to use the Euler-Lagrange equation; this is usually brought up in the case of the Brachistochrone problem, or the Principle of Least Action \footnote{For background, see The Feynman Lectures, vol II chapter 19}. In these types of problems, it appears that the path of integration through space from endpoints a to b is instead replaced by the whole frequency domain from -$\infty$ to +$\infty$.

@@

where $\int e^{j \omega t} d\omega$ is the inverse fourier transform.

$$W = \frac{1}{\eta_0} \int_{-\infty}^{+\infty}{(n_r(\omega))\ |F(\omega)|^2}\ d\omega$$\\

Where $n(\omega)$ is a complex refractive index per Debye relaxation equation, $n_r()$ is the real part of the same, everything else is a real constant. Set up a functional with a lagrange multiplier\\

$$\xi = -e_x(z,t) + \lambda W$$\\

Take the functional ("variational") gradient with respect to $F(\omega)$.

$$ \nabla \xi = -\frac{1}{\sqrt{2\pi}} \left(e^{- j (\omega/c_0)n(\omega)z}\right)^\star \  e^{-j\omega T} + \lambda...$$

At this point, you could choose to perform a second integral over the time-domain Greens function solution to the harmonic oscillator\cite{Complex2020}, eq. 4 and 15,

$$x(t) = \int_{-\infty}^{+\infty}{G(t,t')\ \frac{q\cdot e_x(z,t'))}{m}\ } dt'$$\\

(eq. 15, for the underdamped case)

$$G(t,t') = UnitStep(t-t') \frac{1}{\sqrt{\gamma^2-\omega_0^2}  }\sinh\left({\sqrt{\gamma^2-\omega_0^2}}\ (t-t')\right)$$

But a much more straightforward

Where * is the complex conjugate.\\

The $\nabla H() = H()^\star$ thing seems to come from Franks, Signal Theory, 1969, page 140 onwards; I cannot make heads or tails of his formulation.

Neither Maxima, SageMath, Mathematica's VariationalD, Matlab's Fundiff, nor Maple's 
functionalDerivative seem to yield any meaningful result for grad xi.\\

Then set

$$x(t) = \int_{-\infty}^{+\infty}{G(t,t')\ \frac{q\cdot e_x(z,t'))}{m}\ } dt'$$\\

Anyhow, then modify the lagrange multiplier,

$$\xi = -x(t) + \lambda W$$\\

And now,

$$\nabla \xi = 0$$

This lagrange multiplier technique has, however, a fundamental limitation (noted by Tufts \cite{Optimum1964}): for reasons that the author does not comprehend, only the pulse energy can be used as a constraint, rather than the value of the peak field. This seems to turn out to be of some importance.

As noted by Pozar, the complex conjugate arises from the concept of "adjointness". Franks.

>>>>

\section{Dispersion Timothy and the Brillouin Precursors}

Uzunoglu \cite{Theoretical2020} mention a handy way to circumvent the lossy-flesh problem: the Brillouin and Sommerfeld precursors\footnotemark, and pulse or wavepacket dispersion in tissues in general. The original text of these papers are in German; translations can be found in Léon Brillouin's 1960 book\cite{Wave1960} on the topic.

@@

@@

\begin{itemize}
	\item The highly dispersive nature of tissue (the variation in speed of light with frequency), which causes a peculiar distortion of short pulses (which necessarily have a broadband spectral content).
	\item The variation in loss (the imaginary part of the complex permittivity and the penetration depth) versus frequency. \wikinote{add image from that mathematica code}
\end{itemize}

A third aspect leads to the formation of the prototypical Brillouin precursor specifically: any sharp 
beginning or end of a tone, or other sharply defined structure in a waveform, includes harmonics beyond the tone's carrier wave up to the rise time\wikinote{add image}.

@@

Based on Yang's inactivation thresholds, for instance, at 9 GHz, assuming that a torso is about 10 penetration depths in radius and the hypothetical Brillouin train drives the virus equally effectively as, implementing a precursor would decrease the field required at the skin to a perfectly practicable 1 kV/m, rather than an air-ionizing 6.6 MV/m ($\frac{1}{\sqrt{10}}=0.316$ versus $e^{-10}=4.5e^{-5}$).

Precursors have been used in practice, by Ong \cite{Detection2003}\footnote{Well, Ong don't really observe the precursor itself directly, it seems.} using simple square waves to detect respiration\footnotemark, Ossberger \cite{Noninvasive2004} for the same purpose; and various techniques for tumor detection.\footnote{TODO: Also another really good paper that I can't find}

While it is clearly possible to produce this abnormal penetration (any sharp edge sufficiently distant from suffices to provoke dispersive distortion), it was not obvious to us that a waveform could be constructed that would be an effective driver for the 10 GHz normal modes in question.

First, if dispersive pulses are too close together, they interfere as they propagate and diminish the precursor thus formed \cite{Dynamical2005}.

As an example, despite the waveform from nonlinear transmission lines\cite{NLTL6275} producing strong wideband components, in cursory simulations they did not appear to satisfy the requirements for usable Brillouin pulses\wikinote{dubious-discuss}\wikinote{citation-needed}. Our working hypothesis is that this is because of destructive interference from the falling edge that immediately follows; but this is mere speculation.\wikinote{dubious-discuss}

Second, and most critically, while the pulse amplitude decays slowly, the total energy contained in such a precursor decays very quickly. The biological ramifications of this are discussed in detail by Adair in \cite{Biophysics2000}.

%temporal Soliton-producing nonlinear transmission lines, and NLTL oscillators constructed therefrom, particularly well suited 
%
%At a depth of 8 cm in Cole-Cole muscle, an 8 GHz resonance, attack pattern alpha appears to achieve 4 picometers of oscillation amplitude, rather than 0.01 picometers with $\sin(t\ \omega_{resonance})$. The sharp edges are at approx. 250 GHz.

Dispersive pulse propagation is easily modelled by a Fourier transform, propagation, and an inverse fourier transform (see Franzen \cite{Wideband1999}, see also \cite{Comments1993} and \cite{Shortrisetime1989}). \footnote{This isn't truly a transient process, of course - the fourier integral is still periodic, wrapping around the window; it's just more transient than a steady-state}.

The loss and refractive index can be obtained from a four term Cole-Cole model\cite{dielectric1996}\cite{gabriel1996compilation}. We used parameters from the it.is tissue database \cite{Tissue2018a}.

@@

@@

\begin{autem}
	autem: 
Reflection from tissue interfaces?

@@

\footnote{Progress was hindered by the number of conventions for Fourier normalization. It's all well and good for the signal processing specialists to make all sorts of excuses about brevity, but when crossing fields it's pretty annoying for such a fundamental concept to be so loosely defined. On the same topic, can we finally settle the i vs. j convention, and the choice of reduced units in molecular dynamics?}

@@

We apologize in advance for the desecration of mathematics in this section.

The most promising analytical technique explored was a slight modification of a formulation for optimal-energy pulses, largely supplied by Prof. Costas Sarris in a yet-unpublished personal communication, based on Zhu, Hum, Costas formulation in \cite{Microwave2012b}, previously discussed by Pozar \cite{Waveform2003}.

@@

Here, rather than taking the derivative with respect to a small change in a variable, we take the derivative with respect to a small change in the function itself - a very useful technique \footnote{that the author was not aware of!} which imposes little to no limitation on the form of the solution. For some intuition, the definition of the first variation and the Gateaux derivative may be of some use.\footnote{The complex conjugate in Sarris' formulation, $\nabla H() = H()^\star$, seems to come from \cite{Signal1969}, page 140 onwards; I cannot make heads or tails of his formulation. Pozar note that this complex conjugate arises from the concept of "adjointness".}

A common setup for such a problem is to use the Euler-Lagrange equation; this is usually brought up in the case of the Brachistochrone problem, or the Principle of Least Action \footnote{For background, see The Feynman Lectures, vol II chapter 19}. In these types of problems, it appears that the path of integration through space from endpoints a to b can be replaced by the whole frequency domain from -$\infty$ to +$\infty$.

@@

where $\int e^{j \omega t} d\omega$ is the inverse fourier transform. At this point, you could choose to perform a second integral over the time-domain Greens function solution to the harmonic oscillator\cite{Complex2020}, or substitute

\footnote{Many symbolic math tools have support for functional calculus. However, neither Maxima, SageMath, Mathematica's VariationalD, Matlab's Fundiff, nor Maple's functionalDerivative seem to yield any meaningful result for grad xi.}\\

%$$W = \frac{1}{\eta_0} \int_{-\infty}^{+\infty}{(n_r(\omega))\ |F(\omega)|^2}\ d\omega$$\\
%
%Where $n(\omega)$ is a complex refractive index per Debye relaxation equation, $n_r()$ is the real part of the same, everything else is a real constant. Set up a functional with a lagrange multiplier\\
%
%$$\xi = -e_x(z,t) + \lambda W$$\\
%
%Take the functional ("variational") gradient with respect to $F(\omega)$.
%
%$$ \nabla \xi = -\frac{1}{\sqrt{2\pi}} \left(e^{- j (\omega/c_0)n(\omega)z}\right)^\star \  e^{-j\omega T} + \lambda...$$
%
%At this point, you could choose to perform a second integral over the time-domain Greens function solution to the harmonic oscillator\cite{Complex2020}, eq. 4 and 15, 
%
%$$x(t) = \int_{-\infty}^{+\infty}{G(t,t')\ \frac{q\cdot e_x(z,t'))}{m}\ } dt'$$\\
%
%(eq. 15, for the underdamped case)
%
%$$G(t,t') = UnitStep(t-t') \frac{1}{\sqrt{\gamma^2-\omega_0^2}  }\sinh\left({\sqrt{\gamma^2-\omega_0^2}}\ (t-t')\right)$$
%
%But a much more straightforward 
%
%Where * is the complex conjugate.\\

%
%Then set
%
%$$x(t) = \int_{-\infty}^{+\infty}{G(t,t')\ \frac{q\cdot e_x(z,t'))}{m}\ } dt'$$\\
%
%Anyhow, then modify the lagrange multiplier,
%
%$$\xi = -x(t) + \lambda W$$\\
%
%And now,
%
%$$\nabla \xi = 0$$

This lagrange multiplier technique has, however, a fundamental limitation (noted by Tufts \cite{Optimum1964}): for reasons that the author does not yet comprehend, only the pulse energy can be used as a constraint, rather than the value of the peak field. This seems to turn out to be of some importance.

### 2021-03-01 documents/experiment_1_data.tex
Tersenating again

\begin{table}[h!]
	\centering
	\begin{tabular}{ |c|c|c| } 
		\hline
		Exposure & kilocounts per 10s integration & Note \# \\
		\hline
		Autoclaved phage (0.2 mL) & 30 & \\ 
		Autoclaved p. (50 uL) & 32 & Run 8\\ 
		\hline
		Phage culture 'blank' & 8 & \\ 
		Water & 1 & \\
		\hline
		\multicolumn{3}{|c|}{\ntilde 10 or 20 us  } \\
		\hline
		 & \\ 
		\hline
		\multicolumn{3}{|c|}{Sweep from 6 to 12 GHz, average approximately 150 V/m, 10 or 20 us CW pulses} \\
		\hline
		\multicolumn{3}{|c|}{\ntilde 0.05 mm, \ntilde } \\
		\hline
		\multicolumn{3}{|c|}{\ntilde 0.2 mm channel, 0.2 uL/s, various pulse counts and termination} \\
		\hline 
		 Blank    & 8 & Li test 1\\
		 +           & 8 & Li test 1\\
 		 +           & 7 & Li test 1\\
 		 +           & 7 & Li test 1\\
 		 + \& shaper & 7 & Li test 1\\
 		 \hline
		\multicolumn{3}{|c|}{\ntilde 0.05 mm channel, \ntilde 0.75 MV/m, 0.2 uL/s, various pulse counts and termination} \\
		\hline
		Sham & 5 & Smaller \\
		+ & 6 & Smaller \\ 
		+ & 6 & Smaller  \\ 
		\hline
		\multicolumn{3}{|c|}{\ntilde 0.05 mm channel, \ntilde 0.75 MV/m, 0.2 uL/s, basic buffer, but pH may not have been 10.6} \\
		\hline
		+ & 11 & Weakened \\
		+ (no term) & 11 & Weakened \\ 
		Sham & 12 & Weakened  \\ 
		\hline
	\end{tabular}
\end{table}

Data from testing on bacteriophage. Data rounded (0.5 and above up) to the nearest thousand photon counts per 10 seconds. %thoroughly characterized.

Background ('blank', in notes) readings measure the amount of DNA present in the phage culture tube before any treatment.

Repeatability between test conditions was not terrific. The source of the \ntilde 5000 count variation between some equivalent samples is not known. However, in no case did any tube approach the 22 to 30 kcounts of the autoclave.

Stock culture had titer $3\times 10^9$ T4r.

The fluor-meter printed a constant stream of results, all within the random error between capture windows $\sigma=0.5$ kcounts. Only after most of the data was taken was it realized how dumb this was.

Results aggregated over the course of several months.

In all cases, a thermal effect was implausible; the total energy in the pulse train precluded heating above the capsid melting temperature.\cite{Effects1951}

In an early iteration of the experiment, the firmware and data analysis scripts randomly generated blinded sham and autoclave assignments and automatically treated the samples accordingly. It also printed blinded templates for pipetting. This was done partly in the hopes that massively oversampling the microwave absorption spectrometer across the range of physical tolerances would highlight any resonances.\cite{first2000}

However, as spectrometry was abandoned, this non-deterministic property caused significant issues while troubleshooting the fluorescence detection system, and \st{expedience} \st{\textit{laziness}} prompted us to disable this system, violating best practices.

A thin layer of silicone conformal coating was applied to the 0.2 mm waveguide to avoid the copper ion artifact, but this could not be done on the 0.05 mm waveguide.

>>>>

\subsection{Data from testing on bacteriophage T4}
\begin{table}[h!]
	\centering
	\begin{tabular}{ |c|c|c| } 
				\hline
		\multicolumn{3}{|c|}{\textbf{Calibration trials} } \\
		\hline
		\hline
		Exposure & kilocounts per 10s integration & Note \# \\
		\hline
		Autoclaved phage (0.2 mL) & 30 & \\ 
		Autoclaved p. (50 uL) & 32 & Run 8\\ 
		\hline
		Phage culture 'blank' & 8 & \\ 
		Water & 1 & \\

		\hline
\multicolumn{3}{|c|}{\textbf{Pulsed CW trials}} \\
\hline

		\hline
		\multicolumn{3}{|c|}{\textbf{NF capillary}, 6 to 12 GHz, 10 or 20 us CW pulses} \\
		\hline
		Sham   & 12 & Run 8 t4,5...\\
		+ & 11 & Run 8 t4,5...\\
		+  & 12 & Run 8 t4,5...\\
		+  & 12 & Run 8 t4,5...\\
		\hline
		\multicolumn{3}{|c|}{\textbf{NF capillary}, 6 to 12 GHz, phage degraded with basic buffer} \\
		\hline
		Blank (0 min) & 14 & Run 9\\
		Blank (\ntilde 15 min) & 17 & Run 9\\
		Sham (\ntilde 15 min) & 5 & Run 9\\
		Blank (\ntilde 4 hr) & 24 & Run 9\\
		\hline
		+  & 14 & Run 11\\
		Sham & 17 & Run 11\\
		\hline
		\multicolumn{3}{|c|}{\textbf{NF capillary}, 6 to 12 GHz, 10 or 20 us CW pulses} \\
		\hline
		Sham & 12 & Run 12\\
		+    & 10 & Run 12\\
		\hline
		\multicolumn{3}{|c|}{\textbf{0.2 mm channel}, 6 to 12 GHz, 10 or 20 us CW pulses} \\
		\hline
		 Sham & 12 & CPWG micro-\\
 		 +    & 10 & CPWG micro-\\

		\hline
		\multicolumn{3}{|c|}{\textbf{Nanosecond pulsed trials} } \\
		\hline

 		\hline
		\multicolumn{3}{|c|}{\ntilde 0.2 mm channel, 0.2 uL/s, various pulse counts and termination} \\
		\hline 
		 Blank    & 8 & Li test 1\\
		 +           & 8 & Li test 1\\
 		 +           & 7 & Li test 1\\
 		 +           & 7 & Li test 1\\
 		 + \& shaper & 7 & Li test 1\\
 		 \hline

		\multicolumn{3}{|c|}{\ntilde 0.05 mm channel, 0.2 uL/s, various pulse counts and termination} \\
		\hline
		Sham & 5 & Smaller \\
		+ & 6 & Smaller \\ 
		+ & 6 & Smaller  \\ 
		\hline

		\multicolumn{3}{|c|}{\ntilde 0.05 mm channel, 0.2 uL/s, basic buffer, tainted as pH may not have been 10.6} \\
		\hline
		Sham & 12 & Weakened \\ 
		Blank & 11 & Weakened  \\ 
		Blank & 12 & Weakened  \\ 
		Blank & 9 & Weakened  \\ 
		+ & 11 & Weakened \\
		+ (no term) & 11 & Weakened \\ 
		+ & 9 & Weakened \\ 
		\hline
	\end{tabular}
\end{table}

Original data for this table can be inspected at \ghfile{firmware/eppenwolf/pulse_1.pnw}.

See notes on data collection in main body.
